---
title: "Beads in string"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
Created by <p style="color:blue">*Budhaditya Basu*.</p>

*****

Open RStudio, select `File/New File/R Script` and start writing your script with the help of this tutorial.

# Loading Packages

```{r, echo = TRUE}
library(ggplot2)
```

# Importing data

The tab delimited data frame 'Panther_CC.txt' is assigned to a vector df.

```{r}
df <- read.csv("Panther_CC.txt", sep = "\t", header = TRUE)#The columns are tab separated, and contains header
```

To check the first few data, we can use `head()` function.


```{r}
head(df)
```

# Plotting


To plot the data, we use `ggplot()` function. Define the X and Y axis within aes().

```{r}
ggplot(data = df, aes(x = Status, y = Cellular.Component, fill = FDR,
                    size= Fold_enrichment)) #Fill the color based on FDR value & based on fold enrichment point size will differ
```

# Choose the shape of points

Here we are going to use a filled circle (shape = 21). Shape 19 is solid circle, shape 20 is smaller solid circle, which I don't want to use. The shape size from 21 to 25 can be colored and filled with different colors.

```{r}
p  <- ggplot(data = df, aes(x = Status, y = Cellular.Component, 
                            fill = FDR, size= Fold_enrichment))+
  geom_point(shape=21)
p
```

# Help about geom_points
To get more details of point shapes, you can use `help(point)`.

<br><br><br>
# Change theme and fill color
```{r}
p + theme_bw() + #Change theme to black and white
  scale_fill_gradient2(low="darkcyan",mid ="red", high="blue")# Define fill color
```

# Remove X and Y axis title 
```{r}
p + theme_bw() +
  scale_fill_gradient2(low="darkcyan",mid ="red", high="blue")+
  xlab("")+ ylab("")
```

**The above plot needs to be modified as following:** <br>
1. Change the order of X-axis element i.e., Upregulated first and then Downregulated. <br>
2. Rotate Upregulated and Downregulated text 45 degree. <br>
3. Provide a title and sub title of the plot. <br>
4. Change the legend title. I want to change from 'Fold_enrichment' to 'Fold Enrichment'.

# Rotate the X-axis element
```{r}
p + theme_bw() +
  scale_fill_gradient2(low="darkcyan",mid ="red", high="blue")+
  xlab("")+ ylab("")+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+ # Rotate X-axis element
  scale_x_discrete(limits=c("Upregulated","Downregulated"))+ #Change the order 
  labs(title = "Gene Ontology",
       subtitle = "Cellular Component")+ # Add title and subtitle
  labs(size="Fold Enrichment", fill="FDR") #change legend title and order
```

**********

This plot can be found in the following article:

**Pax6 Modulates Intra-Retinal Axon Guidance and Fasciculation of Retinal Ganglion Cells during Retinogenesis** <br>
<i class="fas fa-link"></i> [nature.com/articles](https://doi.org/10.1038/s41598-020-72828-4)
