---
title: "Violin plot"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
Created by <p style="color:blue">*Budhaditya Basu*.</p>

*****



Open RStudio, select `File/New File/R Script` and start writing your script with the help of this tutorial.

# Loading Packages

```{r, echo = TRUE}
library(ggpubr)
```

# Importing data

The tab delimited data frame 'axon.txt' is assigned to a vector df.


```{r}
df <- read.csv("axon.txt", sep = "\t", header = TRUE) 
#The columns are tab separated, and contains header
```

To check the first few data, we can use `head()` function.


```{r}
head(df)
```

# Plotting


To plot the data, we use `ggplot()` function. Define the X and Y axis within aes()


```{r}
ggplot(data = df, aes(x=Retina, y=Diameter))
```

Let's add the code for the type of graph we need. We are going to plot a violin plot. So the code will look like the following.


```{r}
ggplot(data = df, aes(x=Retina, y=Diameter))+
  geom_violin()
```

# Addition of layers
## Add box plot

To add a box plot within the violin plot:

```{r}
ggplot(data = df, aes(x=Retina, y=Diameter))+
  geom_violin()+
  geom_boxplot(width = 0.1)
```

Change violin plot colors by groups within the column 'Retina' 

```{r}
ggplot(data = df, aes(x=Retina, y=Diameter, fill = Retina))+
  geom_violin()+
  geom_boxplot(width = 0.1)
```

## Add jitters
If we want to see individual data points along with the previous plot, add jitters to it. 
```{r}
p <- ggplot(data = df, aes(x=Retina, y=Diameter, fill = Retina))+
     geom_violin()+
     geom_boxplot(width = 0.1)+
     geom_jitter(position=position_jitter(0.1),
              color="red", size=0.7, alpha=0.5) # 0.1 : degree of jitter in x direction

p
```

## Change theme and remove legend

```{r}
p + 
  theme_classic()+
  theme(legend.position = "none")
```

Change the order of groups in the X axis and give defined colors to it

```{r}
p +
  theme_classic()+
  theme(legend.position = "none")+
  scale_x_discrete(limits=c("WT OD", "KO OD" ,"WT Periphery", "KO periphery")) + 
  scale_fill_manual(values=c("gray87", "gray87", "lightblue", "lightblue"))
```

## Add title

```{r}
p +
  theme_classic()+
  theme(legend.position = "none")+
  scale_x_discrete(limits=c("WT OD", "KO OD" ,"WT Periphery", "KO periphery")) + 
  scale_fill_manual(values=c("gray87", "gray87", "lightblue", "lightblue"))+
  ggtitle("Violin plot") +
  xlab("")+ ylab("Diameter in micrometer")
```

**********

This plot can be found in the following article:

**Pax6 Modulates Intra-Retinal Axon Guidance and Fasciculation of Retinal Ganglion Cells during Retinogenesis** <br>
<i class="fas fa-link"></i> [nature.com/articles](https://doi.org/10.1038/s41598-020-72828-4)

